---
title: "R In-Class Exercise - Functions"
author: "Your Name Here"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#first load the libraries we'll need
library(tidyverse)
library(nycflights13) # <- make sure to install this package 
library(tidycensus)
library(janitor)
library(writexl)
library(here)
options(tigris_class = "sf")

```

# Functions

1) Create a function that will return the sum of 2 numbers.

```{r}
1+1
addtwonumbers <- function(firstnumber, secondnumber) {
  result <- firstnumber + secondnumber
  return(result)
}

addtwonumbers(34, 35)
```

2) Create a function that will return the mean of a list of numbers fed to it. (Remember that in our case we're using the word "list" for easier understanding, but we're talking about a "vector" of numbers in R-speak.)

```{r}
makemymean <- function(mynumbers) {
  result <- mean(mynumbers)
  return(result)
}

makemymean(c(20,30,40))

--
## the way i did it
  
mean_numbers <- function(numbers){
  result <- mean(numbers)
  return(result)
}

mean_numbers(c(1, 2, 3, 4))

```

3) Now, expand on the function you made above, and instead of just returning the mean number alone, have it return the sentence:
*"The mean is ___"*. (Where the blank is the mean.)

```{r}
## the way prof did it

numberlist <- c(11, 21, 15, 83)

resultingmean <- makemymean(numberlist)

paste0(" the mean is! ", resultingmean)

mean_with_text <- function(numberlist){
  resultingmean <- makemymean(numberlist)
  finaltext <- paste0(" the mean is! ", resultingmean)
  return(finaltext)
}

mean_with_text(c(1, 19, 5, 77))


```

For these next two questions, we'll use the `flights` dataset from the package `nycflights13`. It contains records of individual airline flights departing at New York's three metro airports, JFK, LaGuardia (LGA) and Newark (EWR). Run the following chunk to load the data. (Make sure you've actually installed the package, of course.)

```{r}
#load the data to use for the rest of the assignment questions
flights <- nycflights13::flights

head(flights)
```

4) Using the `flights` dataset, create a function that filters the data frame to only include records from a single originating airport (column name `origin`). In other words, it should let you enter an origin airport and return all the records from just that airport.

```{r}
## from class

airportname <- "JFK"

flights %>% 
  filter(origin == airportname)

filter_flights_byorigin_aiporrt <- function(airportname) {
 result <- flights %>% 
  filter(origin == airportname)
  return(result)
}

filter_flights_byorigin_aiporrt("JFK")

---

allfrom_jfk <- function(filter == "origin")){
  result <- paste0(" The mean is ", " ", mean(numbers))
  return(result)
}

```

5) Similarly, using the `flights` dataset, create a function that:

-   filters the data frame to only include records from a single originating airport (column name `origin`), and then

-   aggregates (i.e. counts) the results by airline (`carrier`) to show how many flights each airline has from that airport.

```{r}

airportname <- "JFK"
flights %>% 
  filter(origin == airportname) %>% 
  count(carrier)

top_airline_foroneairport <- function(airportname) {
 
 result <- flights %>% 
  filter(origin == airportname) %>% 
  count(carrier, sort = TRUE) %>% 
   head(1) %>% 
   select(carrier)
 return(result)
}

top_airline_foroneairport("JFK")
```
```{r}

```

